### 2020-11-20 MSDS6371Project.Rmd
add rmd file

---
title: "MSDS6371Project"
author: "Allen Miller & Mingyang Nick YU"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Read in Training Data and initial inspection:
```{r}
#Load required libraries
library(tidyverse)
library(ggplot2)
library(olsrr)
#Load in Dataset
#Call original data amesHouse
amesHouse = read.csv("/Users/mingyang/Desktop/SMU/StatisticalFoundation_Fall2020/MSDS6371Project/train.csv",header = TRUE)
#observe data
head(amesHouse)
str(amesHouse)
#look at neighborhood
amesHouse$Neighborhood
```

#### Analysis 1:
```{r}
#extract variables we're interested for Analysis 1
data1 = amesHouse %>% select(SalePrice,GrLivArea,Neighborhood)%>%
  filter(Neighborhood=="NAmes"|Neighborhood=="Edwards"|Neighborhood=="BrkSide")
str(data1)
#383 observations selected after filtering neighborhood of "NAmes", "Edwards" and "BrkSide"
#Convert Neighborhood to factors
data1$Neighborhood = as.factor(data1$Neighborhood)
#Plot and observe the relationship between SalePrice and GrLivArea
data1 %>% ggplot(aes(x=GrLivArea,y=SalePrice))+
  geom_point()+ggtitle("Sale Price vs. Square Foot of Living area")+
  xlab("Square Foot of Living Area")+
  ylab("Sale Price")
#there are some outliers that may not have come from the same population of interest
#build model to identify the outliers
model1 = lm(SalePrice~GrLivArea,data=data1)
summary(model1)
#Look at residual plots and cook's distance
plot(model1)
cooksDistance = data.frame(distance = cooks.distance(model1))
cooksDistance%>%filter(distance>=1)
ols_plot_cooksd_bar(model1)
ols_plot_resid_stand(model1)
#ols_plot_resid_lev(model1)
# Observation 339 has cook's distance larger than 5.6, Observation 131's cook's D is larger than 1 so it may due to the face it is a unique case.
# Observation 169 and 190 has standarlized residual greater than 4, these two observations can also potentially not coming from the same population of interest
# Since our sample is sufficiently large, deleting these four outliers won't make a huge difference.
# Call these new dataset without outliers data2
data2 = data1[-c(131,169,190,339),]
str(data2)
#build new model to double check
model2 = lm(SalePrice~GrLivArea,data=data2)
summary(model2)
plot(model2)
ols_plot_cooksd_bar(model2)
ols_plot_resid_stand(model2)
#Cook's Distance within roughly 0.1, residual within 3, decent amount of observations beyond 2 (within 5%)
#Assumptions met, move on to include categorical variables
model3 = lm(SalePrice~GrLivArea+Neighborhood+GrLivArea*Neighborhood,data=data2)
summary(model3)
plot(model3)
ols_plot_cooksd_bar(model3)
ols_plot_resid_stand(model3)
#Assumptions are met, model 3 has higher Adjusted R-squared compared to model2

```

### 2020-11-21 MSDS6371Project.Rmd
updates to analysis question 2

@@
 library(olsrr)
+library(GGally)
+library(caret)
+library(MASS)
+library(leaps)
 #Load in Dataset
@@
 #extract variables we're interested for Analysis 1
-data1 = amesHouse %>% select(SalePrice,GrLivArea,Neighborhood)%>%
+data1 = amesHouse %>% dplyr::select(SalePrice,GrLivArea,Neighborhood)%>%
   filter(Neighborhood=="NAmes"|Neighborhood=="Edwards"|Neighborhood=="BrkSide")
@@
 ```
+
+#### Analysis 2 - Variable converting & Inspecting relationship:
+```{r}
+# convert certain columns we need into factor in order to use multi regression model
+cols.to.factor = c("MSSubClass","MSZoning","Street","Alley","LotShape","LandContour","Utilities",
+                   "LotConfig","LandSlope","Neighborhood","Condition1","Condition2","BldgType",
+                   "HouseStyle","OverallQual","OverallCond","RoofStyle","RoofMatl","Exterior1st",
+                   "Exterior2nd","MasVnrType","ExterQual","ExterCond","Foundation","BsmtQual",
+                   "BsmtCond","BsmtExposure","BsmtFinType1","BsmtFinType2","Heating","HeatingQC",
+                   "CentralAir","Electrical","KitchenQual","Functional","FireplaceQu","GarageType",
+                   "GarageFinish","GarageQual","GarageCond","PavedDrive","PoolQC","Fence","MiscFeature",
+                   "SaleType","SaleCondition","PoolArea","MoSold","YrSold","Fireplaces","FullBath")
+amesHouse[cols.to.factor] = lapply(amesHouse[cols.to.factor],factor)
+str(amesHouse)
+
+#Now exploring relationships between continuous variables
+amesHouse %>% dplyr::select(SalePrice,LotFrontage,LotArea,YearBuilt,YearRemodAdd,MasVnrArea,BsmtFinSF1,BsmtFinSF2)%>%
+  ggpairs()
+#Try Logrithmic on SalePrice
+amesHouse$lSalePrice = log(amesHouse$SalePrice)
+#explore relationship between log SalePrice
+amesHouse %>% dplyr::select(lSalePrice,LotFrontage,LotArea,YearBuilt,YearRemodAdd,MasVnrArea,BsmtFinSF1,BsmtFinSF2)%>%
+  ggpairs()
+#There seem to have a slight improvement between YearBuilt and YearRemodAdd
+
+#Move forward with next batch of exploration
+amesHouse %>% dplyr::select(SalePrice,BsmtUnfSF,TotalBsmtSF,X1stFlrSF,X2ndFlrSF,LowQualFinSF,GrLivArea) %>%
+  ggpairs()
+amesHouse %>% dplyr::select(lSalePrice,BsmtUnfSF,TotalBsmtSF,X1stFlrSF,X2ndFlrSF,LowQualFinSF,GrLivArea) %>%
+  ggpairs()
+#There seem to have a good linear relationship between lSalePrice and TotalBsmtSF, X1stFlrSF, X2ndFlrSF,GrLivArea
+#correlation looks better with lSalePrice comparing to SalePrice
+
+#Move forward with next batch of exploration
+amesHouse %>% dplyr::select(lSalePrice,BsmtFullBath,BsmtHalfBath,HalfBath,BedroomAbvGr,KitchenAbvGr,
+                     TotRmsAbvGrd)%>%
+  ggpairs()
+
+#The only variable seem to have good correlation with lSalePrice is FullBath,TotRmsAbvGrd.
+#???If we should change BsmtFullBath,BsmtHalfBath,HalfBath,BedroomAbvGr,KitchenAbvGr to factors???
+#Keep going
+amesHouse %>% dplyr::select(lSalePrice,GarageYrBlt,GarageCars,GarageArea) %>% ggpairs()
+#The above variables all seem to have a good linear correlation with lSalePrice
+#Keep going
+amesHouse %>% dplyr::select(lSalePrice,WoodDeckSF,OpenPorchSF,EnclosedPorch,X3SsnPorch,ScreenPorch,MiscVal)%>%
+  ggpairs()
+#linear relationships don't seem particularly strong except WoodDeckSF, OpenPorchSF
+
+```
+### Select key variables that can be used as predictor - based on results above
+```{r}
+summary(amesHouse)
+#LotFrontage has too many missing values, yet it only has 0.356 correlation with lSalePrice, so not using this variable
+#Alley has too many missing values, so delete column Alley
+#LowQualFinSF has close to no correlaion with lSalePrice, delete
+#BsmtHalfBath has close to 0 correlation,delete
+#FireplaceQu too much missing data, delete
+#X3SsnPorch too little correlation, delete
+#PoolQC too much missing data, delete
+#MiscFeature too much missing data, delete
+#MiscVal too little correlation, delete
+#Consder adding: Exterior1nd, Exterior2nd
+#Selected variables make sense to predict lSalePrice
+amesHouse$PoolYN = ifelse(is.na(amesHouse$PoolQC),"NO","YES")
+amesHouse$PoolYN = as.factor(amesHouse$PoolYN)
+variables.used = amesHouse %>% 
+  dplyr::select(Id,MSSubClass,MSZoning,LotArea,LotConfig,Neighborhood,HouseStyle,
+         OverallQual,OverallCond,YearBuilt,YearRemodAdd,MasVnrArea,ExterQual,
+         ExterCond,Foundation,BsmtQual,BsmtCond,TotalBsmtSF,
+         Heating,CentralAir,X1stFlrSF,GrLivArea,FullBath,KitchenQual,TotRmsAbvGrd,
+         Fireplaces,GarageType,GarageCars,PoolYN,MoSold,YrSold,lSalePrice)
+str(variables.used)
+summary(variables.used)
+#Adjusting MasVnrArea
+variables.used$MasVnrArea = ifelse(is.na(variables.used$MasVnrArea),0,variables.used$MasVnrArea)
+variables.used$BsmtQual = as.character(variables.used$BsmtQual)
+variables.used$BsmtQual = ifelse(is.na(variables.used$BsmtQual),"NONE",variables.used$BsmtQual)
+variables.used$BsmtQual = as.factor(variables.used$BsmtQual)
+#Adjusting BsmtCond
+variables.used$BsmtCond = as.character(variables.used$BsmtCond)
+variables.used$BsmtCond = ifelse(is.na(variables.used$BsmtCond),"NONE",variables.used$BsmtCond)
+variables.used$BsmtCond = as.factor(variables.used$BsmtCond)
+#Adjusting GarageType
+variables.used$GarageType = as.character(variables.used$GarageType)
+variables.used$GarageType = ifelse(is.na(variables.used$GarageType),"NONE",variables.used$GarageType)
+variables.used$GarageType = as.factor(variables.used$GarageType)
+
+temp = variables.used %>% drop_na()
+```
+
+### Stepwise Selection
+```{r}
+set.seed(11)
+train.control <- trainControl(method = "cv", number = 10)
+step.model = train(lSalePrice~.-Id, data=variables.used,
+                   method="lmStepAIC",
+                   trControl = train.control,
+                   trace=FALSE)
+step.model$results
+step.model$finalModel
+summary(step.model$finalModel)
+```

### 2020-11-21 MSDS6371Project.Rmd
update with stepwise output

@@
 summary(step.model$finalModel)
+print(step.model)
+exp(0.1494448)
 ```
+
+### Load in Test Data 
+```{r}
+#Load in test.csv
+testData = read.csv("/Users/mingyang/Desktop/SMU/StatisticalFoundation_Fall2020/MSDS6371Project/test.csv",header = TRUE)
+#Convert variables we need from N/A to NONE
+testData$MasVnrArea = ifelse(is.na(testData$MasVnrArea),0,testData$MasVnrArea)
+testData$TotalBsmtSF = ifelse(is.na(testData$TotalBsmtSF),0,testData$TotalBsmtSF)
+testData$GarageCars = ifelse(is.na(testData$GarageCars),0,testData$GarageCars)
+testData$MSZoning = ifelse(is.na(testData$MSZoning),"RL",testData$MSZoning)
+testData$KitchenQual = ifelse(is.na(testData$KitchenQual),"TA",testData$KitchenQual)
+testData$MSSubClass = ifelse(testData$MSSubClass==150,180,testData$MSSubClass)
+testData$FullBath = ifelse(testData$FullBath==4,3,testData$FullBath)
+testData$Fireplaces = ifelse(testData$Fireplaces==4,3,testData$Fireplaces)
+
+testData$BsmtQual = ifelse(is.na(testData$BsmtQual),"NONE",testData$BsmtQual)
+testData$BsmtCond = ifelse(is.na(testData$BsmtCond),"NONE",testData$BsmtCond)
+testData$GarageType = ifelse(is.na(testData$GarageType),"NONE",testData$GarageType)
+#Convert some variables we need into factor
+testData[cols.to.factor] = lapply(testData[cols.to.factor],factor)
+#create variable PoolYN
+testData$PoolYN = ifelse(is.na(testData$PoolQC),"NO","YES")
+testData$PoolYN = as.factor(testData$PoolYN)
+#select variables needed for testing
+testVariable = testData %>% 
+  dplyr::select(Id,MSSubClass,MSZoning,LotArea,LotConfig,Neighborhood,HouseStyle,
+         OverallQual,OverallCond,YearBuilt,YearRemodAdd,MasVnrArea,ExterQual,
+         ExterCond,Foundation,BsmtQual,BsmtCond,TotalBsmtSF,
+         Heating,CentralAir,X1stFlrSF,GrLivArea,FullBath,KitchenQual,TotRmsAbvGrd,
+         Fireplaces,GarageType,GarageCars,PoolYN,MoSold,YrSold)
+
+```
+
+### run stepwise test and export result to csv file - Kaggle Score 0.15372
+```{r}
+testVariable$lSalePrice = predict(step.model,testVariable)
+testVariable$SalePrice = exp(testVariable$lSalePrice)
+result.stepwise = testVariable %>% dplyr::select(Id,SalePrice)
+write.csv(result.stepwise,"stepwise_Prediction_Miller_YU.csv",row.names = FALSE)
+```

### 2020-11-21 MSDS6371Project.Rmd
adding forward and backward selection

@@
 
-temp = variables.used %>% drop_na()
 ```
@@
 exp(0.1494448)
+ols_plot_resid_stand(step.model$finalModel)
+ols_plot_cooksd_bar(step.model$finalModel)
 ```
@@
 ```
+
+### Forward Selection - Kaggle Score 0.15432
+```{r}
+set.seed(12)
+#create forward fit model
+
+#write a program to pretict 100 different combination to test forward selection model
+iterations = 100
+splitPerc = 0.8
+total_RMSE = 0
+for(i in 1:iterations){
+  trainIndices = sample(1:dim(variables.used)[1],round(splitPerc * dim(variables.used)[1]))
+  train = variables.used[trainIndices,]
+  test = variables.used[-trainIndices,]
+  forward_fit = lm(lSalePrice~.-Id,data=train)
+  forward.model = ols_step_forward_aic(forward_fit,penter=0.15)
+  prediction = predict(forward.model$model,test)
+  squared_MSPE = mean((test$lSalePrice - prediction)^2)
+  temp_RMSE = sqrt(squared_MSPE)
+  total_RMSE = total_RMSE+temp_RMSE
+}
+#CV Press=1.003755
+exp(total_RMSE/iterations)
+#Build model with full dataset for fitting
+forward_fit = lm(lSalePrice~.-Id,data=variables.used)
+forward.model = ols_step_forward_aic(forward_fit,penter=0.15)
+forward.model$model
+forward.model
+ols_plot_resid_stand(forward.model$model)
+ols_plot_cooksd_bar(forward.model$model)
+plot(forward.model$model)
+
+testVariable$lSalePriceForward = predict(forward.model$model,testVariable)
+testVariable$SalePrice = exp(testVariable$lSalePriceForward)
+result.forward = testVariable %>% dplyr::select(Id,SalePrice)
+write.csv(result.forward,"Forward_Prediction_Miller_YU_2.csv",row.names = FALSE)
+
+```
+
+### Backward Selection - Kaggle Score 0.15432
+```{r}
+set.seed(15)
+#create forward fit model
+
+#write a program to pretict 100 different combination to test forward selection model
+iterations = 100
+splitPerc = 0.8
+total_RMSE = 0
+for(i in 1:iterations){
+  trainIndices = sample(1:dim(variables.used)[1],round(splitPerc * dim(variables.used)[1]))
+  train = variables.used[trainIndices,]
+  test = variables.used[-trainIndices,]
+  backward_fit = lm(lSalePrice~.-Id,data=train)
+  backward.model = ols_step_backward_aic(backward_fit)
+  prediction = predict(backward.model$model,test)
+  squared_MSPE = mean((test$lSalePrice - prediction)^2)
+  temp_RMSE = sqrt(squared_MSPE)
+  total_RMSE = total_RMSE+temp_RMSE
+}
+#CV Press = 1
+exp(total_RMSE/iterations)
+#Build model with full dataset for fitting
+backward_fit = lm(lSalePrice~.-Id,data=variables.used)
+backward.model = ols_step_backward_aic(backward_fit)
+backward.model$model
+backward.model
+ols_plot_resid_stand(backward.model$model)
+ols_plot_cooksd_bar(backward.model$model)
+plot(backward.model$model)
+
+testVariable$lSalePriceBackward = predict(backward.model$model,testVariable)
+testVariable$SalePrice = exp(testVariable$lSalePriceBackward)
+result.backward = testVariable %>% dplyr::select(Id,SalePrice)
+write.csv(result.backward,"Backward_Prediction_Miller_YU.csv",row.names = FALSE)
+```
+
+### Backward Selection - use p-value - Kaggle 0.15505
+```{r}
+set.seed(16)
+#create forward fit model
+
+#write a program to pretict 100 different combination to test forward selection model
+iterations = 100
+splitPerc = 0.8
+total_RMSE = 0
+for(i in 1:iterations){
+  trainIndices = sample(1:dim(variables.used)[1],round(splitPerc * dim(variables.used)[1]))
+  train = variables.used[trainIndices,]
+  test = variables.used[-trainIndices,]
+  backward_fit = lm(lSalePrice~.-Id,data=train)
+  backward.model = ols_step_backward_p(backward_fit,prem=0.1,details=FALSE)
+  prediction = predict(backward.model$model,test)
+  squared_MSPE = mean((test$lSalePrice - prediction)^2)
+  temp_RMSE = sqrt(squared_MSPE)
+  total_RMSE = total_RMSE+temp_RMSE
+}
+#CV Press = 1
+exp(total_RMSE/iterations)
+#Build model with full dataset for fitting
+backward_fit = lm(lSalePrice~.-Id,data=variables.used)
+backward.model = ols_step_backward_p(backward_fit,prem=0.1,details=FALSE)
+backward.model$model
+backward.model
+ols_plot_resid_stand(backward.model$model)
+ols_plot_cooksd_bar(backward.model$model)
+plot(backward.model$model)
+
+testVariable$lSalePriceBackward = predict(backward.model$model,testVariable)
+testVariable$SalePrice = exp(testVariable$lSalePriceBackward)
+result.backward = testVariable %>% dplyr::select(Id,SalePrice)
+write.csv(result.backward,"Backward_Prediction_usingP_Miller_YU.csv",row.names = FALSE)
+```

### 2020-11-22 MSDS6371Project.Rmd
updates

@@
 #Move forward with next batch of exploration
+amesHouse %>% dplyr::select(SalePrice,BsmtFullBath,BsmtHalfBath,HalfBath,BedroomAbvGr,KitchenAbvGr,
+                     TotRmsAbvGrd)%>%
+  ggpairs()
 amesHouse %>% dplyr::select(lSalePrice,BsmtFullBath,BsmtHalfBath,HalfBath,BedroomAbvGr,KitchenAbvGr,
@@
 #Keep going
+amesHouse %>% dplyr::select(SalePrice,GarageYrBlt,GarageCars,GarageArea) %>% ggpairs()
 amesHouse %>% dplyr::select(lSalePrice,GarageYrBlt,GarageCars,GarageArea) %>% ggpairs()
@@
 #Keep going
+amesHouse %>% dplyr::select(SalePrice,WoodDeckSF,OpenPorchSF,EnclosedPorch,X3SsnPorch,ScreenPorch,MiscVal)%>%
+  ggpairs()
 amesHouse %>% dplyr::select(lSalePrice,WoodDeckSF,OpenPorchSF,EnclosedPorch,X3SsnPorch,ScreenPorch,MiscVal)%>%
@@
 #linear relationships don't seem particularly strong except WoodDeckSF, OpenPorchSF
+#Overall, lSalePrice seem to improve the correlation with our important predictors a bit. So we will proceed
+#with lSalePrice Analysis
 
@@
 }
+#total RMSE = 0.00374748... RMSE = 839.29 under original SalePrice scale
+total_RMSE/iterations
 #CV Press=1.003755
@@
 write.csv(result.forward,"Forward_Prediction_Miller_YU_2.csv",row.names = FALSE)
+
+#Try Mass Library Forward model
+forward.model2 = stepAIC(forward_fit,direction="forward")
+forward.model2$model
 
@@
 write.csv(result.backward,"Backward_Prediction_Miller_YU.csv",row.names = FALSE)
+
+#Try Mass Library Forward model - Same result 0.15432
+backward.model2 = stepAIC(backward_fit,direction="backward")
+backward.model2
+
+#Testing with Mass Library
+testVariable$lSalePriceBackward = predict(backward.model2,testVariable)
+testVariable$SalePrice = exp(testVariable$lSalePriceBackward)
+result.backward = testVariable %>% dplyr::select(Id,SalePrice)
+write.csv(result.backward,"Backward_Prediction_2_Miller_YU.csv",row.names = FALSE)
 ```

### 2020-11-22 MSDS6371Project.Rmd
update rmd

--- 2020-11-21 MSDS6371Project.Rmd
@@
 print(step.model)
+# RMSE = 0.1494448 by 10 fold internal cross-validation
 exp(0.1494448)
+ols_plot_resid_stand(step.model$finalModel)
+ols_plot_cooksd_bar(step.model$finalModel)
 ```
@@
 ```
+
+### Forward Selection - Kaggle Score 0.15432
+```{r}
+set.seed(115)
+#create forward fit model
+
+#write a program to pretict 5 different combination to test forward selection model
+iterations = 5
+splitPerc = 0.9
+total_RMSE = 0
+for(i in 1:iterations){
+  print(i)
+  trainIndices = sample(1:dim(variables.used)[1],round(splitPerc * dim(variables.used)[1]))
+  train = variables.used[trainIndices,]
+  test = variables.used[-trainIndices,]
+  forward_fit = lm(lSalePrice~.-Id,data=train)
+  forward.model = ols_step_forward_aic(forward_fit,penter=0.15)
+  prediction = predict(forward.model$model,test)
+  squared_MSPE = mean((test$lSalePrice - prediction)^2)
+  temp_RMSE = sqrt(squared_MSPE)
+  total_RMSE = total_RMSE+temp_RMSE
+}
+#total RMSE = 0.1235978... RMSE = 839.29 under original SalePrice scale
+total_RMSE/iterations
+
+#Build model with full dataset for fitting
+forward_fit = lm(lSalePrice~.-Id,data=variables.used)
+forward.model = ols_step_forward_aic(forward_fit,penter=0.15)
+forward.model$model
+forward.model
+ols_plot_resid_stand(forward.model$model)
+ols_plot_cooksd_bar(forward.model$model)
+plot(forward.model$model)
+
+testVariable$lSalePriceForward = predict(forward.model$model,testVariable)
+testVariable$SalePrice = exp(testVariable$lSalePriceForward)
+result.forward = testVariable %>% dplyr::select(Id,SalePrice)
+write.csv(result.forward,"Forward_Prediction_Miller_YU_2.csv",row.names = FALSE)
+
+#Try Mass Library Forward model
+forward.model2 = stepAIC(forward_fit,direction="forward")
+forward.model2$model
+
+```
+
+### Backward Selection - Kaggle Score 0.15432
+```{r}
+set.seed(112)
+#create forward fit model
+
+#write a program to pretict 5 different combination to test forward selection model
+iterations = 5
+splitPerc = 0.95
+total_RMSE = 0
+for(i in 1:iterations){
+  print(i)
+  trainIndices = sample(1:dim(variables.used)[1],round(splitPerc * dim(variables.used)[1]))
+  train = variables.used[trainIndices,]
+  test = variables.used[-trainIndices,]
+  backward_fit = lm(lSalePrice~.-Id,data=train)
+  backward.model = ols_step_backward_aic(backward_fit)
+  prediction = predict(backward.model$model,test)
+  squared_MSPE = mean((test$lSalePrice - prediction)^2)
+  temp_RMSE = sqrt(squared_MSPE)
+  total_RMSE = total_RMSE+temp_RMSE
+}
+#CV Press Estimate = 0.1481379
+total_RMSE/iterations
+
+#Build model with full dataset for fitting
+backward_fit = lm(lSalePrice~.-Id,data=variables.used)
+backward.model = ols_step_backward_aic(backward_fit)
+backward.model$model
+backward.model
+ols_plot_resid_stand(backward.model$model)
+ols_plot_cooksd_bar(backward.model$model)
+plot(backward.model$model)
+
+testVariable$lSalePriceBackward = predict(backward.model$model,testVariable)
+testVariable$SalePrice = exp(testVariable$lSalePriceBackward)
+result.backward = testVariable %>% dplyr::select(Id,SalePrice)
+write.csv(result.backward,"Backward_Prediction_Miller_YU.csv",row.names = FALSE)
+
+#Try Mass Library Forward model - Same result 0.15432
+backward.model2 = stepAIC(backward_fit,direction="backward")
+backward.model2
+
+#Testing with Mass Library
+testVariable$lSalePriceBackward = predict(backward.model2,testVariable)
+testVariable$SalePrice = exp(testVariable$lSalePriceBackward)
+result.backward = testVariable %>% dplyr::select(Id,SalePrice)
+write.csv(result.backward,"Backward_Prediction_2_Miller_YU.csv",row.names = FALSE)
+```
+
+### Backward Selection - use p-value - Kaggle 0.15505
+```{r}
+set.seed(16)
+#create forward fit model
+
+#write a program to pretict 20 different combination to test forward selection model
+iterations = 20
+splitPerc = 0.95
+total_RMSE = 0
+for(i in 1:iterations){
+  print(i)
+  trainIndices = sample(1:dim(variables.used)[1],round(splitPerc * dim(variables.used)[1]))
+  train = variables.used[trainIndices,]
+  test = variables.used[-trainIndices,]
+  backward_fit = lm(lSalePrice~.-Id,data=train)
+  backward.model = ols_step_backward_p(backward_fit,prem=0.1,details=FALSE)
+  prediction = predict(backward.model$model,test)
+  squared_MSPE = mean((test$lSalePrice - prediction)^2)
+  temp_RMSE = sqrt(squared_MSPE)
+  total_RMSE = total_RMSE+temp_RMSE
+}
+#CV Press estimate = 0.1306067
+total_RMSE/iterations
+#Build model with full dataset for fitting
+backward_fit = lm(lSalePrice~.-Id,data=variables.used)
+backward.model = ols_step_backward_p(backward_fit,prem=0.1,details=FALSE)
+backward.model$model
+backward.model
+ols_plot_resid_stand(backward.model$model)
+ols_plot_cooksd_bar(backward.model$model)
+plot(backward.model$model)
+
+testVariable$lSalePriceBackward = predict(backward.model$model,testVariable)
+testVariable$SalePrice = exp(testVariable$lSalePriceBackward)
+result.backward = testVariable %>% dplyr::select(Id,SalePrice)
+write.csv(result.backward,"Backward_Prediction_usingP_Miller_YU.csv",row.names = FALSE)
+```

### 2020-11-23 MSDS6371Project.Rmd
adding custom model

@@
 
-### run stepwise test and export result to csv file - Kaggle Score 0.15372
+### run stepwise test and export result to csv file - Kaggle Score 0.15372, CV Press = 0.1494, Adjusted R2 = 0.8975
 ```{r}
@@
 
-### Forward Selection - Kaggle Score 0.15432
+### Forward Selection - Kaggle Score 0.15432, CV Press estimate = 0.14199, Adjusted R2 = 0.89575
 ```{r}
-set.seed(115)
+set.seed(116)
 #create forward fit model
 
-#write a program to pretict 5 different combination to test forward selection model
+#write a program to generate 5 different combination to test forward selection model
 iterations = 5
@@
 }
-#total RMSE = 0.1235978... RMSE = 839.29 under original SalePrice scale
+#total RMSE = 0.1235978... 
 total_RMSE/iterations
@@
 
-### Backward Selection - Kaggle Score 0.15432
+### Backward Selection - Kaggle Score 0.15432, CV-Press = 0.148, adjusted R2 = 0.89575
 ```{r}
@@
 ```
+
+### Custom Model - Kaggle Score = 0.14773, CV-Press = 0.1649, Adjusted R2 = 0.9207
+```{r}
+#reload data
+amesHouse = read.csv("/Users/mingyang/Desktop/SMU/StatisticalFoundation_Fall2020/MSDS6371Project/train.csv",header = TRUE)
+summary(amesHouse)
+library(mice) #Load mice library to analyze the pattern of missing data
+md.pattern(amesHouse)
+#imputing missing data using predictive mean matching
+tempData <- mice(amesHouse,m=1,maxit=50,meth='pmm',seed=20)
+amesHouse = complete(tempData,1)
+summary(amesHouse)
+
+#Convert certain variables into factors
+cols.to.factor = c("MSSubClass","MSZoning","Street","Alley","LotShape","LandContour","Utilities",
+                   "LotConfig","LandSlope","Neighborhood","Condition1","Condition2","BldgType",
+                   "HouseStyle","OverallQual","OverallCond","RoofStyle","RoofMatl","Exterior1st",
+                   "Exterior2nd","MasVnrType","ExterQual","ExterCond","Foundation","BsmtQual",
+                   "BsmtCond","BsmtExposure","BsmtFinType1","BsmtFinType2","Heating","HeatingQC",
+                   "CentralAir","Electrical","KitchenQual","Functional","FireplaceQu","GarageType",
+                   "GarageFinish","GarageQual","GarageCond","PavedDrive","PoolQC","Fence","MiscFeature",
+                   "SaleType","SaleCondition","PoolArea","MoSold","YrSold","Fireplaces","FullBath")
+amesHouse[cols.to.factor] = lapply(amesHouse[cols.to.factor],factor)
+
+amesHouse$PoolYN = ifelse(is.na(amesHouse$PoolQC),"NO","YES")
+amesHouse$PoolYN = as.factor(amesHouse$PoolYN)
+#First test with first few variables
+models1 = regsubsets(SalePrice~MSSubClass+MSZoning+LotFrontage+LotArea+Street+Alley+LotShape, data = amesHouse, nvmax = 10)
+summary(models1)
+#Choose the optimal model
+res.sum <- summary(models1)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# As we can see among the variables selected MSSubClass, MSZoning,LotArea,Alley has variables that are significant
+
+#Next set of variables
+models2 = regsubsets(SalePrice~LandContour+Utilities+LotConfig+LandSlope+Neighborhood+Condition1+Condition2, data = amesHouse, nvmax = 15)
+summary(models2)
+#Choose the optimal model
+res.sum <- summary(models2)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# LotConfig, Neighborhood, Condition2 has most variables that can be significant
+
+#Next set of variables
+models3 = regsubsets(SalePrice~BldgType+HouseStyle+OverallQual+OverallCond+YearBuilt, data = amesHouse, nvmax = 10)
+summary(models3)
+#Choose the optimal model
+res.sum <- summary(models3)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# BldgType, Neighborhood, OverallQual, YearBuilt has most variables that can be significant
+
+#Next set of variables
+models4 = regsubsets(SalePrice~YearRemodAdd+RoofStyle+RoofMatl+Exterior1st+Exterior2nd, data = amesHouse, nvmax = 10)
+summary(models4)
+#Choose the optimal model
+res.sum <- summary(models4)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# YearRemodAdd, RoofStyleHip, RoofMatl, Exterior1st, Exterior2nd has most variables that can be significant
+
+#Next set of variables
+models5 = regsubsets(SalePrice~MasVnrType+MasVnrArea+ExterQual+ExterCond+Foundation, data = amesHouse, nvmax = 10)
+summary(models5)
+#Choose the optimal model
+res.sum <- summary(models5)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# MasVnrType, MasVnrArea, ExterQual, Foundation, have most variables that can be significant
+
+#Next set of variables all basement section
+models6 = regsubsets(SalePrice~BsmtQual+BsmtCond+BsmtExposure+BsmtFinType1+BsmtFinSF1+
+                       BsmtFinType2+BsmtFinSF2+BsmtUnfSF+TotalBsmtSF, data = amesHouse, nvmax = 15)
+summary(models6)
+#Choose the optimal model
+res.sum <- summary(models6)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# BsmtQual, BsmtFinType1, TotalBsmtSF have most variables that can be significant
+
+#Next set of variables  - Utilities
+models7 = regsubsets(SalePrice~Heating+HeatingQC+CentralAir+Electrical, data = amesHouse, nvmax = 10)
+summary(models7)
+#Choose the optimal model
+res.sum <- summary(models7)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# Heating, HeatingQC, CentralAir,Elecrical have most variables that can be significant
+
+#Next set of variables  - Square foots
+models8 = regsubsets(SalePrice~X1stFlrSF+X2ndFlrSF+LowQualFinSF+GrLivArea+BsmtFullBath+BsmtHalfBath+FullBath+HalfBath, data = amesHouse, nvmax = 10)
+summary(models8)
+#Choose the optimal model
+res.sum <- summary(models8)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# X1stFlrSF, X2ndFlrSF, LowQualFinSF,GrLivArea,BsmtFullBath,BsmtHalfBath,FullBath1,HalfBath have most variables that can be significant
+
+#Next set of variables  - Square foots
+models9 = regsubsets(SalePrice~BedroomAbvGr+KitchenAbvGr+KitchenQual+TotRmsAbvGrd+Functional, data = amesHouse, nvmax = 10)
+summary(models9)
+#Choose the optimal model
+res.sum <- summary(models9)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# BedroomAbvGr, KitchenAbvGr, KitchenQual,TotRmsAbvGrd have most variables that can be significant
+
+#Next set of variables  
+models10 = regsubsets(SalePrice~Fireplaces+FireplaceQu+GarageType+GarageYrBlt+
+                        GarageFinish+GarageCars+GarageArea+GarageQual+GarageCond, data = amesHouse, nvmax = 10)
+summary(models10)
+#Choose the optimal model
+res.sum <- summary(models10)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# Fireplaces, FireplaceQu, GarageFinish,GarageCars,GarageArea have most variables that can be significant
+
+#Next set of variables  
+models11 = regsubsets(SalePrice~PavedDrive+WoodDeckSF+OpenPorchSF+EnclosedPorch+
+                        X3SsnPorch+ScreenPorch, data = amesHouse, nvmax = 10)
+summary(models11)
+#Choose the optimal model
+res.sum <- summary(models11)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# PavedDriveY, WoodDeckSF, OpenPorchSF,X3SsnPorch,ScreenPorch have most variables that can be significant
+
+#Next set of variables  
+models12 = regsubsets(SalePrice~PoolYN+Fence+MiscVal+MoSold+YrSold, data = amesHouse, nvmax = 10)
+summary(models12)
+#Choose the optimal model
+res.sum <- summary(models12)
+data.frame(
+  Adj.R2 = which.max(res.sum$adjr2),
+  CP = which.min(res.sum$cp),
+  BIC = which.min(res.sum$bic)
+)
+# PoolYN, Fence, MoSold,YrSold have most variables that can be significant
+
+#Select Variables
+custom.var = amesHouse %>% 
+  dplyr::select(Id,MSSubClass, MSZoning,LotArea,LotConfig, Neighborhood, Condition2,
+                BldgType, Neighborhood, OverallQual, YearBuilt,YearRemodAdd, RoofStyle, RoofMatl, Exterior1st, Exterior2nd,
+                MasVnrType, MasVnrArea, ExterQual, Foundation, BsmtQual, BsmtFinType1, TotalBsmtSF,
+                Heating, HeatingQC, CentralAir,Electrical,
+                X1stFlrSF, X2ndFlrSF, LowQualFinSF,GrLivArea,BsmtFullBath,BsmtHalfBath,FullBath,HalfBath,
+                BedroomAbvGr, KitchenAbvGr, KitchenQual,TotRmsAbvGrd,
+                Fireplaces,GarageCars,GarageArea,
+                PavedDrive, WoodDeckSF, OpenPorchSF,X3SsnPorch,ScreenPorch,
+                PoolYN, MoSold,YrSold,
+                SalePrice)
+custom.var$SalePrice = log(custom.var$SalePrice)
+custom.var$GrLivArea = log(custom.var$GrLivArea)
+
+#Adjusting NA factors
+custom.var$MasVnrType = as.character(custom.var$MasVnrType)
+custom.var$MasVnrType = ifelse(is.na(custom.var$MasVnrType),"NONE",custom.var$MasVnrType)
+custom.var$MasVnrType = as.factor(custom.var$MasVnrType)
+
+custom.var$BsmtQual = as.character(custom.var$BsmtQual)
+custom.var$BsmtQual = ifelse(is.na(custom.var$BsmtQual),"NONE",custom.var$BsmtQual)
+custom.var$BsmtQual = as.factor(custom.var$BsmtQual)
+
+custom.var$BsmtFinType1 = as.character(custom.var$BsmtFinType1)
+custom.var$BsmtFinType1 = ifelse(is.na(custom.var$BsmtFinType1),"Unf",custom.var$BsmtFinType1)
+custom.var$BsmtFinType1 = as.factor(custom.var$BsmtFinType1)
+
+custom.var$Electrical = as.character(custom.var$Electrical)
+custom.var$Electrical = ifelse(is.na(custom.var$Electrical),"SBrkr",custom.var$Electrical)
+custom.var$Electrical = as.factor(custom.var$Electrical)
+
+#Model Training
+set.seed(11)
+train.control <- trainControl(method = "cv", number = 10)
+custom.model = train(SalePrice~.-Id, data=custom.var,
+                   method="lmStepAIC",
+                   trControl = train.control,
+                   trace=FALSE)
+custom.model$results
+custom.model$finalModel
+summary(custom.model$finalModel)
+print(custom.model)
+# RMSE = 0.1494448 by 10 fold internal cross-validation
+
+ols_plot_resid_stand(custom.model$finalModel)
+ols_plot_cooksd_bar(custom.model$finalModel)
+plot(custom.model$finalModel)
+#Load in testing data and fill in the missing data in the same manner
+testData = read.csv("/Users/mingyang/Desktop/SMU/StatisticalFoundation_Fall2020/MSDS6371Project/test.csv",header = TRUE)
+testData$MSSubClass = ifelse(testData$MSSubClass==150,180,testData$MSSubClass)
+testData$FullBath = ifelse(testData$FullBath==4,3,testData$FullBath)
+testData$Fireplaces = ifelse(testData$Fireplaces==4,3,testData$Fireplaces)
+testData$MasVnrArea = ifelse(is.na(testData$MasVnrArea),0,testData$MasVnrArea)
+testData$TotalBsmtSF = ifelse(is.na(testData$TotalBsmtSF),0,testData$TotalBsmtSF)
+testData$BsmtFullBath = ifelse(is.na(testData$BsmtFullBath),0,testData$BsmtFullBath)
+testData$BsmtHalfBath = ifelse(is.na(testData$BsmtHalfBath),0,testData$BsmtHalfBath)
+testData$GarageCars = ifelse(is.na(testData$GarageCars),0,testData$GarageCars)
+testData$GarageArea = ifelse(is.na(testData$GarageArea),0,testData$GarageArea)
+testData[cols.to.factor] = lapply(testData[cols.to.factor],factor)
+
+testData$PoolYN = ifelse(is.na(testData$PoolQC),"NO","YES")
+testData$PoolYN = as.factor(testData$PoolYN)
+summary(testData)
+#Testing test set
+testData1 = testData %>% 
+  dplyr::select(Id,MSSubClass, MSZoning,LotArea,LotConfig, Neighborhood, Condition2,
+                BldgType, Neighborhood, OverallQual, YearBuilt,YearRemodAdd, RoofStyle, RoofMatl, Exterior1st, Exterior2nd,
+                MasVnrType, MasVnrArea, ExterQual, Foundation, BsmtQual, BsmtFinType1, TotalBsmtSF,
+                Heating, HeatingQC, CentralAir,Electrical,
+                X1stFlrSF, X2ndFlrSF, LowQualFinSF,GrLivArea,BsmtFullBath,BsmtHalfBath,FullBath,HalfBath,
+                BedroomAbvGr, KitchenAbvGr, KitchenQual,TotRmsAbvGrd,
+                Fireplaces,GarageCars,GarageArea,
+                PavedDrive, WoodDeckSF, OpenPorchSF,X3SsnPorch,ScreenPorch,
+                PoolYN, MoSold,YrSold)
+testData1$GrLivArea = log(testData1$GrLivArea)
+summary(testData1)
+#Adjusting NA factors
+testData1$MSZoning = as.character(testData1$MSZoning)
+testData1$MSZoning = ifelse(is.na(testData1$MSZoning),"RL",testData1$MSZoning)
+testData1$MSZoning = as.factor(testData1$MSZoning)
+
+testData1$Exterior1st = as.character(testData1$Exterior1st)
+testData1$Exterior1st = ifelse(is.na(testData1$Exterior1st),"Other",testData1$Exterior1st)
+testData1$Exterior1st = as.factor(testData1$Exterior1st)
+
+testData1$Exterior2nd = as.character(testData1$Exterior2nd)
+testData1$Exterior2nd = ifelse(is.na(testData1$Exterior2nd),"Other",testData1$Exterior2nd)
+testData1$Exterior2nd = as.factor(testData1$Exterior2nd)
+
+testData1$BsmtQual = as.character(testData1$BsmtQual)
+testData1$BsmtQual = ifelse(is.na(testData1$BsmtQual),"TA",testData1$BsmtQual)
+testData1$BsmtQual = as.factor(testData1$BsmtQual)
+
+testData1$BsmtFinType1 = as.character(testData1$BsmtFinType1)
+testData1$BsmtFinType1 = ifelse(is.na(testData1$BsmtFinType1),"Unf",testData1$BsmtFinType1)
+testData1$BsmtFinType1 = as.factor(testData1$BsmtFinType1)
+
+testData1$KitchenQual = as.character(testData1$KitchenQual)
+testData1$KitchenQual = ifelse(is.na(testData1$KitchenQual),"TA",testData1$KitchenQual)
+testData1$KitchenQual = as.factor(testData1$KitchenQual)
+
+testData1$MasVnrType = as.character(testData1$MasVnrType)
+testData1$MasVnrType = ifelse(is.na(testData1$MasVnrType),"None",testData1$MasVnrType)
+testData1$MasVnrType = as.factor(testData1$MasVnrType)
+
+testData1$Exterior1st = as.character(testData1$Exterior1st)
+testData1$Exterior1st = ifelse(testData1$Exterior1st=="Other","VinylSd",testData1$Exterior1st)
+testData1$Exterior1st = as.factor(testData1$Exterior1st)
+
+testData1$SalePrice = predict(custom.model,testData1)
+testData1$SalePrice = exp(testData1$SalePrice)
+custom.result = testData1 %>% dplyr::select(Id,SalePrice)
+write.csv(custom.result,"custom_model_Miller_YU.csv",row.names = FALSE)
+```

### 2020-11-23 MSDS6371Project.Rmd
adjust rmd

@@
 
-#write a program to generate 5 different combination to test forward selection model
+#write a program to generate 5 different combination to Cross-Validate Forward Selection Model
 iterations = 5

### 2020-11-25 MSDS6371Project.Rmd
update MSDS report

@@
 str(data2)
+data2 %>% ggplot(aes(x=GrLivArea,y=SalePrice))+
+  geom_point()+ggtitle("Sale Price vs. Square Foot of Living area")+
+  xlab("Square Foot of Living Area")+
+  ylab("Sale Price")
 #build new model to double check
@@
 model3 = lm(SalePrice~GrLivArea+Neighborhood+GrLivArea*Neighborhood,data=data2)
+data2 %>% ggplot(aes(x=GrLivArea,y=SalePrice,color=Neighborhood))+
+  geom_point()+ggtitle("Sale Price vs. Square Foot of Living area")+
+  xlab("Square Foot of Living Area")+
+  ylab("Sale Price")
 summary(model3)
@@
 ols_plot_resid_stand(model3)
+anova(model3)
+confint(model3)
 #Assumptions are met, model 3 has higher Adjusted R-squared compared to model2
+# Change reference to Edwards to get CI for slope
+data2$Neighborhood = relevel(data2$Neighborhood,ref="Edwards")
+model3 = lm(SalePrice~GrLivArea+Neighborhood+GrLivArea*Neighborhood,data=data2)
+summary(model3)
+confint(model3)
+
+# Change reference to NAmes to get CI for slope
+data2$Neighborhood = relevel(data2$Neighborhood,ref="NAmes")
+model3 = lm(SalePrice~GrLivArea+Neighborhood+GrLivArea*Neighborhood,data=data2)
+summary(model3)
+confint(model3)
 
